Supporting online material for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-08-15 Documents/supplementary_text.tex
Working on supplement.

\usepackage{amsmath}
\usepackage[pdfborder={0 0 0}]{hyperref}
\usepackage[numbers,round,sort&compress]{natbib}
\usepackage{hypernat}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{microtype}
\usepackage[final]{pdfpages}

@@

@@

Several sources provided information to parameterize our model at the
country level. Demographic data, including population growth rate
\cite{WorldBankpg}, death rate
\cite{World_Development_Indicators2013-ee}, and number of people aged
15--49 years \cite{The_World_Bank2016-fd} were obtained from the World
Bank. Longitudinal HIV incidence and prevalence estimates for the
15--49 age group, spanning from as early as 1990 to 2015, was
primarily derived from the AIDSinfo database produced by UNAIDS
\cite{Unaids2016-an}. Other sources included UNAIDS Country Progress
Reports \cite{Unaids2016-am} published from 2012 to 2016 and AIDS Data
Hub \cite{AIDSdatahub-fg}, which compiles data from UNAIDS, UNICEF,
WHO, and the Asian Development Bank. These sources were also used to
inform the initial conditions for the model: the number of people who
have been diagnosed with HIV, who are on ART, and who have viral
suppression or have been retained on treatment for at least 12
months. Where data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  We found sufficient data to
parametrize our mathematical modeling for 127 countries.  See
Supplementary Material 2 (SM2)for full information on data sources for
each country.

\section{Mathematical model}
\begin{figure}
  \label{model_diag}
  \begin{center}
    \input{diagram}
    \caption{Model diagram.}
  \end{center}
\end{figure}

@@

\section{Model fitting}
\label{model_fitting}
We first estimated country-specific rates of HIV transmission for each
of the available historical data point for incidence and
prevalence. The calculation involved a simplified force of infection
that initially assumed no differences in transmission risk among
acute, unsuppressed, suppressed, and AIDS classes due to the general
lack of historical data on diagnosis and treatment. We calculated a
time-independent aggregate transmission rate by taking an exponential
weighted mean of the historical (i.e., time-dependent) transmission
rates for each given country (\ref{transmission_rate}). The aggregate
transmission rate was then combined with estimated HIV transmission
risk by stage of HIV infection \cite{Hollingsworth2008-iy} and
treatment status \cite{Wawer2005-us} to calculate separate
transmission rates for unsuppressed, acute, and viral-suppression
people. These rates were used to simulate model projections through
2035.

@@

\section{Analysis}
We evaluated the global and country-specific impacts of meeting the
UNAIDS targets and vaccine deployment on four outcome measures: HIV
incidence, PLHIV, number of people with AIDS, and AIDS-related deaths
between 2015 and 2035. Calibrating our model to historical data on
prevalence and incidence, we projected the anticipated trajectories of
the outcomes using country-specific transmission rates across the
intervention scenarios outlined below.  Our base-case scenario assumed
maintenance of the status quo: that 2015 levels of the proportion of
PLHIV diagnosed ($p_{D}$), proportion of those diagnosed on ART
($p_{T}$), and proportion of those on treatment who have achieved
viral suppression ($p_{V}$) continue unchanged until 2035.  The
intervention scenarios were (1.) the attainment of the 90--90--90
target by linear increases in $p_{D}$, $p_{T}$, and $p_{V}$ until 2020
followed by maintenance of those levels until 2035, and (2.) linear
increase in the diagnosis and treatment levels to the 90--90--90
targets by 2020, a linear increase from 90--90--90 to the 95--95--95
target by 2030, and maintenance of the 95--95--95 levels to 2035. (For
90--90--90, if any initial level was above 90\%, it was kept constant
going forward so that the intervention did not worsen coverage.
Likewise, for 95--95--95, any initial level above 95\% was kept
constant.)  For vaccination, we assumed deployment of a vaccine with
50\% efficacy in 2020 with coverage increasing linearly from 0\% to
50\% at 2 years post-rollout and maintaining constant coverage after
reaching 70\%. The added benefit of vaccination was considered in the
context of status quo diagnosis and treatment, achieving the UNAIDS
90–90–90 goal by 2020 followed by maintenance, and achieving the
95–95–95 goal by 2030 followed by maintenance.  Projections were
generated through model simulations based on published parameter
distributions (\autoref{model_param}) and country-specific estimated
transmission rate distributions. The parameters were sampled using
latin hypercube sampling \cite{} and the variance reduction technique
of running different interventions with random parameter values
\cite{}.

We calculated partial rank correlation coefficients (PRCC) to measure
the independent effect of each parameter on model projections of HIV
incidence. Furthermore, due to uncertainty surrounding the prospective
vaccine, we obtained model projections using six additional vaccine
scenarios by considering alternative estimates of efficacy (30\%,
70\%), coverage (50\%, 90\%), time to 50\% coverage (2 years, 5
years), and first year of availability (2025).

\section{Results}

\subsection{Medians and distributions for country-specific
  transmission rates}
\label{transmission_rate}
Country-specific transmission rates were calculated based on the
relative contributions of individuals with acute infection,
unsuppressed viral load or suppressed viral load to overall
transmission (\autoref{model_fitting}). The estimates were calibrated
to data on country-level incidence and/or prevalence to account for
variation in historical incidence while accurately reflecting recent
HIV rates.
\includepdf[]{../../Codes/plots/transmission_rate}

@@

Model outcomes were projected until 2035 for each of the 127 countries
with sufficient data on status quo rates of diagnosis and treatment
and on historical incidence or prevalence. The median estimates with
interquartile ranges are presented between 2015 and 2035 for number of
PLHIV, annual incidence (per million population), number of people
with AIDS, and number of HIV-related deaths under the status quo and
five intervention scenarios: vaccination only, attainment of the
UNAIDS 90-90-90 target, 90-90-90 target and vaccination, attainment of
the UNAIDS 95-95-95 target, and 95-95-95 target and vaccination.  The
country-specific projections as well as aggregated projections at the
regional and global levels follow:

@@

\bibliography{SM}

>>>>

\usepackage{amsmath}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage[utf8x]{inputenc}
\usepackage{microtype}
\usepackage[final]{pdfpages}
\usepackage{float}
\usepackage{longtable}
\usepackage[pdfborder={0 0 0}]{hyperref}
\usepackage[numbers,round,sort&compress]{natbib}
\usepackage{hypernat}

@@

\renewcommand{\figurename}{Fig.}

@@

Several sources provided information to parameterize our model at the
country level. Demographic data, including population growth rate
\cite{WorldBankpg}, death rate
\cite{World_Development_Indicators2013-ee}, and number of people aged
15--49 years \cite{The_World_Bank2016-fd} were obtained from the World
Bank. Longitudinal HIV incidence and prevalence estimates for the
15--49 age group, spanning from as early as 1990 to 2015, was
primarily derived from the AIDSinfo database produced by UNAIDS
\cite{Unaids2016-an}. Other sources included UNAIDS Country Progress
Reports \cite{Unaids2016-am} published from 2012 to 2016 and AIDS Data
Hub \cite{AIDSdatahub-fg}, which compiles data from UNAIDS, UNICEF,
WHO, and the Asian Development Bank. These sources were also used to
inform the initial conditions for the model: the number of people who
have been diagnosed with HIV, who are on ART, and who have viral
suppression or have been retained on treatment for at least 12
months. Where data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  We found sufficient data to
parametrize our mathematical modeling for 127 countries.  See
\autoref{data_source_treatment_cascade}--\ref{data_source_regions} for
full information on data sources for each country.

\begin{table}
  \centering
  \tiny
  \input{treatment_cascade.tex}
  \caption{Data sources on the treatment cascade.}
  \label{data_source_treatment_cascade}
\end{table}

\section{Mathematical model}

\begin{figure}
  \centering
  \input{diagram}
  \caption{Model diagram.}
  \label{model_diag}
\end{figure}

@@

\section{Model fitting}
\label{model_fitting}
We first estimated country-specific rates of HIV transmission for each
of the available historical data point for incidence and
prevalence. The calculation involved a simplified force of infection
that initially assumed no differences in transmission risk among
acute, unsuppressed, suppressed, and AIDS classes due to the general
lack of historical data on diagnosis and treatment. We calculated a
time-independent aggregate transmission rate by taking an exponential
weighted mean of the historical (i.e., time-dependent) transmission
rates for each given country (\autoref{transmission_rate}). The
aggregate transmission rate was then combined with estimated HIV
transmission risk by stage of HIV infection
\cite{Hollingsworth2008-iy} and treatment status \cite{Wawer2005-us}
to calculate separate transmission rates for unsuppressed, acute, and
viral-suppression people. These rates were used to simulate model
projections through 2035.

@@

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{../Codes/plots/transmission_rate}
  \caption{Country-specific transmission rates were calculated based
    on the relative contributions of individuals with acute infection,
    unsuppressed viral load or suppressed viral load to overall
    transmission (\autoref{model_fitting}). The estimates were
    calibrated to data on country-level incidence and/or prevalence to
    account for variation in historical incidence while accurately
    reflecting recent HIV rates.}
  \label{transmission_rate}
\end{figure}

\section{Analysis}

We evaluated the global and country-specific impacts of meeting the
UNAIDS targets and vaccine deployment on four outcome measures: HIV
incidence, PLHIV, number of people with AIDS, and AIDS-related deaths
between 2015 and 2035. Calibrating our model to historical data on
prevalence and incidence, we projected the anticipated trajectories of
the outcomes using country-specific transmission rates across the
intervention scenarios outlined below.  Our base-case scenario assumed
maintenance of the status quo: that 2015 levels of the proportion of
PLHIV diagnosed ($p_{D}$), proportion of those diagnosed on ART
($p_{T}$), and proportion of those on treatment who have achieved
viral suppression ($p_{V}$) continue unchanged until 2035.  The
intervention scenarios were (1.) the attainment of the 90--90--90
target by linear increases in $p_{D}$, $p_{T}$, and $p_{V}$ until 2020
followed by maintenance of those levels until 2035, and (2.) linear
increase in the diagnosis and treatment levels to the 90--90--90
targets by 2020, a linear increase from 90--90--90 to the 95--95--95
target by 2030, and maintenance of the 95--95--95 levels to 2035. (For
90--90--90, if any initial level was above 90\%, it was kept constant
going forward so that the intervention did not worsen coverage.
Likewise, for 95--95--95, any initial level above 95\% was kept
constant.)  For vaccination, we assumed deployment of a vaccine with
50\% efficacy in 2020 with coverage increasing linearly from 0\% to
50\% at 2 years post-rollout and maintaining constant coverage after
reaching 70\%. The added benefit of vaccination was considered in the
context of status quo diagnosis and treatment, achieving the UNAIDS
90–90–90 goal by 2020 followed by maintenance, and achieving the
95–95–95 goal by 2030 followed by maintenance.  Projections were
generated through model simulations based on published parameter
distributions (\autoref{model_param}) and country-specific estimated
transmission rate distributions. The parameters were sampled using
latin hypercube sampling \cite{}
and the variance reduction technique of running different
interventions with random parameter values \cite{}.

We calculated partial rank correlation coefficients (PRCC) to measure
the independent effect of each parameter on model projections of HIV
incidence (\autoref{PRCCs}). Furthermore, due to uncertainty
surrounding the prospective vaccine, we obtained model projections
using six additional vaccine scenarios by considering alternative
estimates of efficacy (30\%, 70\%), coverage (50\%, 90\%), time to
50\% coverage (2 years, 5 years), and first year of availability
(2025).

\begin{figure}
  \centering
  \includegraphics{../Codes/plots/prcc.pdf}
  \caption{Parameter sensitivity of model global infections averted
    from 2015 to 2035.}
  \label{PRCCs}
\end{figure}

@@

Model outcomes were projected until 2035 for each of the 127 countries
with sufficient data on status quo rates of diagnosis and treatment
and on historical incidence or prevalence. The median estimates with
interquartile ranges are presented between 2015 and 2035 for number of
PLHIV, annual incidence (per million population), number of people
with AIDS, and number of HIV-related deaths under the status quo and
five intervention scenarios: vaccination only, attainment of the
UNAIDS 90--90--90 target, 90--90--90 target and vaccination, attainment of
the UNAIDS 95--95--95 target, and 95--95--95 target and vaccination.  The
country-specific projections as well as aggregated projections at the
regional and global levels follow:

@@

\bibliography{supplementary_text}

\clearpage
\input{../Codes/plots/effectiveness_all/all}
